More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-06-13 2016-Weak-FH-Bisim/WeakBisim.tex
minor change in todo
changement reu 13/6

\TODO{Id(s) is the identity function on variables of state s}
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\[ \openrule
         {
           (j\mapsto\tau),\symb{True},\symb{Id}(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and for all $\alpha$, $j$, $s$, $\Post$ and $s'$ we have
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \implies \alpha=\tau\land s=s'\land \Post=\symb{Id}(s)
\]
\end{definition}

We then have the following property that justifies the fact that we can abstract away tau transitions from holes.

\begin{property}
There is an open transition from state $s$ to state $s'$ if and only if there is a sequence of open transitions from $s$ to $s'$ where all open transitions except one are of the form:\[ \openrule
         {
           (j\mapsto\tau),\symb{True},\symb{Id}(s)}
         {s \OTarrow {\tau} s}
\]
\end{property}
\TODO{the formal view is not much interesting}

@@

\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},Id(s)}
         {s \OTWeakarrow {\tau} s} \in \WT
 \end{mathpar}
and
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule{
 \openrule
         {
           \overline{\beta},\Pred,\Post}
         {s \OTarrow {\tau} s'} \in \mathcal{T}
\\
 \openrule
         {
           \overline{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
         {s' \OTWeakarrow {\tau} s''} \in \WT
}
{ \openrule
         {
           \vis{\overline{\beta}}\dotcup\overline{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
         {s \OTWeakarrow {\tau} s''} \in \WT
}
 \end{mathpar}

\noindent{\bf Other transitions:}
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule {\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad\qquad
\openrule
         {
           \overline{\beta},\Pred_2,\Post_2  }
         {s_1 \OTarrow {\alpha} s_2} \in \mathcal{T}
\\
\openrule
         {
           \overline{\gamma_3},\Pred_3,({x_k\gets e_k})^{k\in K}    }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}
}
{
\openrule
         {
           \overline{\gamma_1}\dotcup \vis{\overline{\beta}}\dotcup\overline{\gamma_3},
		\Pred,
				({x_k\gets (e_k\subst {\Post_2})\subst{\Post_1}})^{k\in K} } 
         {s \OTWeakarrow {\alpha} s'} \in\WT
}
\end{mathpar}

@@

@@

\TODO{We will discuss this corollary and the title.}
\TODO{Move the following propertiesafter the definition of Weak bisim!}
\begin{corollary}[Weak Composability]
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{corollary}

\subsection{Non-observability and open pNets}
\begin{definition}[Non-observability of silent actions]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \overline{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\symb{True}}\in \overline{\symb{SV}}$ and $\forall i,j\in I\uplus J.\,i\neq j \Rightarrow \alpha_i\neq\tau$
\end{definition}

@@

>>>>

INtiuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
\TODO{Id(s) is the identity function on variables of state s}
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\symb{True},\symb{Id}(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Post=\symb{Id}(s)
\]
\end{enumerate}
\end{definition}

We then can add any number of $\tau$ transitions of the holes before or after any open transition freely. This property justifies the fact that we can abstract away $\tau$ transitions from holes in the definition of a weak open transition.

@@

%\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},Id(s)}
         {s \OTWeakarrow {\tau} s} \in \WT
 \end{mathpar}
and
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule{
 \openrule
         {
           \overline{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
}
{ \openrule
         {
           \vis{\overline{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s''} \in \WT
}
%\inferrule{
% \openrule
%         {
%           \overline{\beta},\Pred,\Post}
%         {s \OTarrow {\tau} s'} \in \mathcal{T}
%\\
% \openrule
%         {
%           \overline{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
%         {s' \OTWeakarrow {\tau} s''} \in \WT
%}
%{ \openrule
%         {
%           \vis{\overline{\beta}}\dotcup\overline{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
%				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
%         {s \OTWeakarrow {\tau} s''} \in \WT
%}
 \end{mathpar}
and
%\noindent{\bf Other transitions:}
\begin{mathpar}
  \mprset {vskip=.5ex}
\inferrule {\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2  }
         {s_1 \OTWeakarrow {\alpha} s_2} \in \WT
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,({x_k\gets e_k})^{k\in K}    }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}
}
{
\openrule
         {
           \overline{\gamma_1}\dotcup \overline{\gamma_2}\dotcup\overline{\gamma_3},
		\Pred,
				({x_k\gets (e_k\subst {\Post_2})\subst{\Post_1}})^{k\in K} } 
         {s \OTWeakarrow {\alpha} s'} \in\WT
}
\end{mathpar}

@@

\TODO{the formal view is not much interesting}

\subsection{Weak FH-bisimulation}

Weak bisimulation can be alternatively defined as a strong bisimulation on the weak open automata, however this would still need a specific definition of strong bisimulation for weau open automata. \TODO{polish}

@@

\subsection{Weak bisimulation for open pNets}
\begin{definition}[Non-observability of silent actions]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \overline{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\symb{True}}\in \overline{\symb{SV}}$ and $\forall i,j\in I\uplus J.\,i\neq j \Rightarrow \alpha_i\neq\tau$
\end{definition}

@@

\begin{property}[Decomposition of weak open transitions] \TODO{TODO}
Let $A$ be the open automaton that is the semantics of a \pNet $P$, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
Suppose we have the following weak open transitions in $\WT$:
\[\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1}
\qquad 
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2   }
         {s_1 \OTWeakarrow {\alpha} s_2}
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,\Post_3   }
         {s_2 \OTWeakarrow {\tau} s_3}
\]
Then there is ag weak  open transition of the following form in $\WT$:
\[\openrule
         {
           \overline{\gamma},\Pred,\Post   }
         {s \OTWeakarrow {\alpha} s_3}
\]
Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
\[
\begin{array}{rcl}
\gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
\Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
\Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
\end{array}\]
\end{property}

\TODO{We will discuss this corollary and the title.}
\begin{corollary}[Weak Composability]
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{corollary}
